# Hello, for some quick context on the project (why I have a 2.5 year-old project in my portfolio)
# This was my most comprehensive project that was a culmination of 3 quarters/1 year of taking R-adjacent classes
# All of the methods used were directly related to structures learned in linear algebra and statistical modeling +machine learning
# This was a collaborative project, but I built all of the code with [Classmate 1] so treat it as everything I can do

---
title: "Spotify Success - Standardized"
author: "[Classmate 1], [Classmate 2], Carson"
date: "2023-12-05"
output: html_document
---

---
title: "INDE 427 - Spotify Project"
author: "[Classmate 1], [Classmate 2], Carson"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) #clears global environment
```

```{r}
## Setting up the data
# The first step is initializing the data. Compared to the downloaded data set from kaggle, we need to address the missing data. For the key variable, we used https://musicstax.com/track/what-was-i-made-for-from-the-motion-picture-barbie/6wf7Yu7cxBSPrRlWeSeK0Q to fill in the missing keys. For the in_shazam_charts variable--we just dropped it (for now).
#Spotify = read.csv("Spotify Success.csv") #initial data set
spotify = read.csv("spotify-2023.csv")
for (i in 1:nrow(spotify)) {
  if(spotify[i, 5] < 1950) {
    spotify[i, 5] = "pre 1950s"
  } else if (1950 <= spotify[i, 5] &&  spotify[i, 5] < 1970) {
    spotify[i, 5] = "1950-1960s"
  } else if (1970 <= spotify[i, 5] && spotify[i, 5] < 1990) {
    spotify[i, 5] = "1970-1980s"
  } else if (1990 <= spotify[i, 5] && spotify[i, 5] < 2010) {
    spotify[i, 5] = "1990s-2000s"
  } else {
    spotify[i, 5] = "post 2010s"
  }
}
for (i in 1:6) {
  spotify[,i] = as.factor(spotify[,i])
} 
for (i in 7:ncol(spotify)) {
  spotify[,i] = as.numeric(spotify[,i])
}
spotify = cbind(spotify[1:6],sapply(spotify[,7:ncol(spotify)], function(x) ifelse(is.na(x), mean(x, na.rm =TRUE), x)))
# spotify = cbind(spotify[1:4],sapply(spotify[,5:ncol(spotify)], function(x) ifelse(is.na(x), median(x, na.rm =TRUE), x)))

x = spotify[,3:(ncol(spotify)-1)] # excluding song name and artist from data
y = spotify[,ncol(spotify)]
y = scale(y)
df_spotify = data.frame(x,y)
names(df_spotify)[ncol(df_spotify)] = c("streams")

## Plotting Points
for (i in 1:ncol(x)) {
  plot(x[,i], y, xlab = names(x)[i], ylab = "streams")
}
```
```{r}
## Setting up Cross Validation
# Base Data
train = sample(nrow(spotify),floor( nrow(spotify) * 7/10) )
spotify_train = df_spotify[train,]
spotify_test = df_spotify[-train,] #
n_folds = 10
N = dim(spotify_test)[1] # sample size
folds_i = sample(rep(1:n_folds, length.out = N))
cv_mse = NULL
```

``` {r}
## Principal Component Analysis - Analyzing Dimensions of Data
require(FactoMineR)
perform_pca = PCA(x[,7:ncol(x)], graph = FALSE, ncp = 10)
require(factoextra )
fviz_screeplot(perform_pca, addlabels = TRUE, ylim = c(0, 50))
# fviz_contrib(spotify_pca, choice = "var", axes = 1, top = 20)
# fviz_contrib(spotify_pca, choice = "var", axes = 2, top = 20)
x_pca = cbind(x[1:6],perform_pca$ind$coord)
spotify_pca = data.frame(x_pca, y)
names(spotify_pca)[7:ncol(spotify_pca)] =   c("PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10", "streams")
# PCA Data
spotify_trainPCA = spotify_pca[train,]
spotify_testPCA = spotify_pca[-train,]
# PCA Cross Validation
perform_pcaCV = PCA(spotify_train[,7:(ncol(x)-1)], graph = FALSE, ncp = 10)
fviz_screeplot(perform_pcaCV, addlabels = TRUE, ylim = c(0, 50)) # Training data leads to similar, yet notably different PCA axis. Analysis was not done on this PCA, but instead the PCA from the full data
# fviz_contrib(spotify_pca, choice = "var", axes = 1, top = 20)
# fviz_contrib(spotify_pca, choice = "var", axes = 2, top = 20)
# x_pcaCV = cbind(x[1:6],perform_pcaCV$ind$coord)
# spotify_pca = data.frame(x_pca, y)
# names(spotify_pca)[7:ncol(spotify_pca)] =   c("PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10", "streams")
```




``` {r}
## LASSO
x_LASSO = cbind(x[1],x[4:5], x[7:8], x[10:ncol(x)])
require(glmnet)
# Conducted LASSO on full data because it was very fincky on what inputs it would accept: may or may not be more correct to conduct on crossvalidation. It would be good practice to do it on training data
fit = glmnet(x, y, family=c("gaussian"))
# Analyze Beta of fit. Variables that appear in earlier columns are more important
# Analyzing Beta you find: in_apple_playlist, in_spotify_playlist, in_spotify_charts, artist_count, energy_., released_year | are the important variable (cutoff of top 6)
# fit$beta[,2]
# fit$beta[,25]
# fit$beta[,27]
# fit$beta[,29]
plot(fit,label = TRUE) # Can remove variables to make picture better
fit_modified = glmnet(x_LASSO,y, family=c("gaussian")) # A LASSO graph w/o x2, x3, x6 and x9 (variables with troublesomely large values that make it hard to view plot)
plot(fit_modified, label = TRUE)

# LASSO Regression
require("ggfortify")
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  modelLASSO <- lm(streams ~ in_apple_playlists+ in_spotify_playlists+ in_spotify_charts+ artist_count+ energy_.+ released_year, data  = spotify_trainCV)
  y_hat <- predict(modelLASSO, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
#summary(modelLASSO)
autoplot(modelLASSO, which = 1:6, ncol = 3, label.size = 3)
```


```{r}
## Linear Regression
#require("ggfortify")
# linearSpotify = lm(streams ~ ., data = spotify_train)
# summary(linearSpotify)
# y_hat = predict(linearSpotify, spotify_test)
# true_y = spotify_test$streams
# cv_mse = mean((true_y - y_hat)^2)
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  model1 <- lm(streams ~., data  = spotify_trainCV)
  y_hat <- predict(model1, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
# summary(model1)
autoplot(model1, which = 1:6, ncol = 3, label.size = 3)

#spotify_lmReduced <- step(model1, direction="backward", test="F")
cv_mse <- NULL
#model1_reduced = streams ~ released_day + in_apple_charts + in_deezer_charts + artist_count + energy_. + in_spotify_charts + released_year + in_apple_playlists + in_spotify_playlists
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  model1_reduced <- lm(streams ~released_day + in_apple_charts + in_deezer_charts + artist_count + energy_. + in_spotify_charts + released_year + in_apple_playlists + in_spotify_playlists, data  = spotify_trainCV)
  y_hat <- predict(model1_reduced, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
# summary(model1_reduced)
autoplot(model1_reduced, which = 1:6, ncol = 3, label.size = 3)
#anova(model1_reduced,model1)

# # Regression on PCA
# lmSpotifyPCA = lm(streams~., data = spotify_trainPCA)
# #summary(lmSpotifyPCA)
# y_hpca = predict(lmSpotifyPCA, spotify_testPCA)
# true_ypca = spotify_testPCA$streams
# cv_msePCA = mean((true_ypca - y_hpca)^2)

cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCVPCA <- spotify_trainPCA[-test_i, ]
  spotify_testCVPCA <- spotify_trainPCA[test_i, ]
  model1_PCA <- lm(streams ~., data  = spotify_trainCVPCA)
  y_hat <- predict(model1_PCA, spotify_testCVPCA)
  true_y <- spotify_testCVPCA$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
# summary(model1_PCA)
autoplot(model1_PCA, which = 1:6, ncol = 3, label.size = 3)

#model1_PCA_reduced = step(model1_PCA, direction="backward", test="F")
#model1_PCA_reduced = streams ~ PC9 + mode + PC5 + artist_count + released_year + PC8 + PC2 + PC4 + PC3 + PC1
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCVPCA <- spotify_trainPCA[-test_i, ]
  spotify_testCVPCA <- spotify_trainPCA[test_i, ]
  model1_PCA_reduced <- lm(streams ~PC9 + mode + PC5 + artist_count + released_year + PC8 + PC2 + PC4 + PC3 + PC1, data  = spotify_trainCVPCA)
  y_hat <- predict(model1_PCA_reduced, spotify_testCVPCA)
  true_y <- spotify_testCVPCA$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
# summary(model1_PCA_reduced)
autoplot(model1_PCA, which = 1:6, ncol = 3, label.size = 3)
```

```{r}
## Random Forest
require(randomForest)
# Default RF with various ntree's, nodesize, and mtry's
# ntree = 100, nodesize = 20, mtry = 5
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 100, nodesize = 20, mtry = 5)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)

# # ntree = 50, nodesize = 20, mtry = 5
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 50, nodesize = 20, mtry = 5)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)
# 
# # ntree = 200, nodesize = 20, mtry = 5
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 200, nodesize = 20, mtry = 5)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)
# 
# # ntree = 200, nodesize = 40, mtry = 5
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 200, nodesize = 40, mtry = 5)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)
# 
# # ntree = 200, nodesize = 30, mtry = 5
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 200, nodesize = 30, mtry = 5)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)

# ntree = 200, nodesize = 20, mtry = 7
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 200, nodesize = 20, mtry = 7)
  y_hat <- predict(spotify_rf, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
mean(cv_mse)
# summary(spotify_rf)
# This seems to be the best parameters for a rf of model1 (all variables)

# ntree = 200, nodesize = 20, mtry = 3
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ ., data = spotify_train, ntree = 200, nodesize = 30, mtry = 3)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)

#Comparing to LASSO and LM_Reduced
# LASSO

# ntree = 200, nodesize = 20, mtry = 3
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ in_apple_playlists+ in_spotify_playlists+ in_spotify_charts+ artist_count+ energy_.+ released_year, data = spotify_train, ntree = 200, nodesize = 20, mtry = 3)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# mean(cv_mse)
# summary(spotify_rf)

# ntree = 200, nodesize = 20, mtry = 2
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  spotify_rf <- randomForest(streams ~ in_apple_playlists+ in_spotify_playlists+ in_spotify_charts+ artist_count+ energy_.+ released_year, data = spotify_train, ntree = 50, nodesize = 20, mtry = 2)
  y_hat <- predict(spotify_rf, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
# summary(spotify_rf)
mean(cv_mse)

# LM_Reduced
# ntree = 200, nodesize = 20, mtry = 2
cv_mse <- NULL
for (k in 1:n_folds) {
  test_i = which(folds_i == k)
  spotify_trainCV <- spotify_train[-test_i, ]
  spotify_testCV <- spotify_train[test_i, ]
  spotify_rf <- randomForest(streams ~ released_day + in_apple_charts + in_deezer_charts + artist_count + energy_. + in_spotify_charts + released_year + in_apple_playlists + in_spotify_playlists, data = spotify_train, ntree = 200, nodesize = 20, mtry = 2)
  y_hat <- predict(spotify_rf, spotify_testCV)
  true_y <- spotify_testCV$streams
  cv_mse[k] <- mean((true_y - y_hat)^2)
}
# summary(spotify_rf)
mean(cv_mse)

# # ntree = 200, nodesize = 20, mtry = 3
# cv_mse <- NULL
# for (k in 1:n_folds) {
#   test_i = which(folds_i == k)
#   spotify_trainCV <- spotify_train[-test_i, ]
#   spotify_testCV <- spotify_train[test_i, ]
#   spotify_rf <- randomForest(streams ~ released_day + in_apple_charts + in_deezer_charts + artist_count + energy_. + in_spotify_charts + released_year + in_apple_playlists + in_spotify_playlists, data = spotify_train, ntree = 200, nodesize = 40, mtry = 3)
#   y_hat <- predict(spotify_rf, spotify_testCV)
#   true_y <- spotify_testCV$streams
#   cv_mse[k] <- mean((true_y - y_hat)^2)
# }
# # summary(spotify_rf)
# mean(cv_mse)

# Conclude that base RF is the best, but LASSO performs decently close whereas the step reduced model is by far the worst.
```
